---
title: 'FDS 3.1 Data Wrangling Exercise 1'
author: "Oswaldo Salazar"
date: "March 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic Data Manipulation

This exercise consisted in manipulating product purchasing data from an electronics store.

### 0: Load Data in RStudio


The original data set was in MS Excel format, so to be able to manipulate it in RStudio the first request was to convert it to CSV and load it.


```{r}
refine_original <- read.csv("~/Documents/Personal Documents/Personal/R/refine_original.csv")
head(refine_original)
```

After loading, I assigned the data frame to a working one, to have the original untouched. Then checked the class of the new data frame.

```{r}
refine_working <- refine_original
class(refine_working)
```

### 1: Clean up brand names

In this part, I standardized the product company name according to the instructions: phillips, akzo, van houten and unilever. To achieve this I used the function grepl to seach for matched to an argument, in this case the company name.

```{r}
refine_working$company[grepl("p", refine_working$company)] <- "phillips"
refine_working$company[grepl("k", refine_working$company)] <- "akzo"
refine_working$company[grepl("t", refine_working$company)] <- "van houten"
refine_working$company[grepl("r", refine_working$company)] <- "unilever"
```
 
 Then checked the commands results:
 
```{r}
head(refine_working)
```

### 2: Separate product code and number

The following instruction asked to separete the product code and product number and put them in separate columns. I used the substr function, passing as arguments the location of the characters needed for each new value.

```{r}
refine_working$product_code <- substr(refine_working$Product.code...number, 1,1)
refine_working$product_number <- substr(refine_working$Product.code...number, 3,4)
head(refine_working)
```

### 3: Add product categories

The next step asked to assign the product category depending on each product code:
* p = Smartphone
* v = TV
* x = Laptop
* q = Tablet

I used a simple ifelse conditional fucntion

```{r}
refine_working$product_category <- ifelse(refine_working$product_code == "p", "Smartphone", 
    ifelse(refine_working$product_code == "v", "TV",
    ifelse(refine_working$product_code == "x", "Laptop", "Tablet")))

head(refine_working)
```

### 4: Add full address for geocoding

Created a new column full_address that concatenates the three address fields (address, city, country), separated by commas.

```{r}
refine_working$full_address <- paste(refine_working$address, refine_working$city, refine_working$country, sep = ", ")

head(refine_working)
```

### 5: Create dummy variables for company and product category

Created dummy binary variables foer each company and product, four new variables for each company and four new variables for each product using ifelse.

```{r}
refine_working$company_phillips <- ifelse(refine_working$company == "phillips", 1, 0)
refine_working$company_akzo <- ifelse(refine_working$company == "akzo", 1, 0)
refine_working$company_van_houten <- ifelse(refine_working$company == "van houten", 1, 0)
refine_working$company_unilever <- ifelse(refine_working$company == "unilever", 1, 0)

refine_working$product_smartphone <- ifelse(refine_working$product_category == "Smartphone", 1, 0)
refine_working$product_tv <- ifelse(refine_working$product_category == "TV", 1, 0)
refine_working$product_laptop <- ifelse(refine_working$product_category == "Laptop", 1, 0)
refine_working$product_tablet <- ifelse(refine_working$product_category == "Tablet", 1, 0)

head(refine_working)
```

Let's take a look a the final working data frame and save it to a CSV.

```{r}
refine_working
write.csv(refine_working, file = "refine_clean.csv")
```